---
title: "Stanley Smith Plant Establishment Trial"
author: "Katie Miller"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
execute:
  message: false
  warning: false
---

```{r}
library(tidyverse) # general usage
library(here) # organization
library(plotly) # interactive plots
library(skimr) # quick glimpses at data
library(naniar) # visualizing missing data
library(magick) # inserting images into plots
library(NatParksPalettes) # one example of color palette package
library(patchwork) # putting plots together
library(GGally)
library(ggeffects)
library(flextable) 
library(car)
library(broom)
library(corrplot)
```


```{r}
# reading in data
#delete "Plant-Trial-main" to render
pet_data <- read.csv(here("data", "plantcodefinal.csv"))

appearance <- read.csv(here("data", "Appearance.csv"))

mortality <- read.csv(here("data", "mortality.csv"))

# cleaning data for analysis
pet_data_clean <-  pet_data %>% 
   filter(irrigation %in% c("high", "low"))
  

sp_labels <- labeller(
  species = c(
    arcto = "Arctostaphylos",
    arte = "Artemisia",
    ceo = "Ceanothus",
    dip = "Diplacus",
    salvia = "Salvia"))


month_labels <- c("May", "Jun", "Jul", "Aug", "Sept", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Aug", "Sept", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr")

selected_breaks <- c(2, 6, 10, 14, 18, 22)
# Define the corresponding labels for these breaks
selected_labels <- month_labels[selected_breaks]

```


```{r}
# initial plot to visualize dataset
ggplot(data = pet_data_clean, aes(x = surv_month, y = growth_rate, color = irrigation)) +
       geom_jitter(na.rm = TRUE) +
  theme_bw()+
  facet_wrap(~species + size, labeller = sp_labels) +
  scale_x_continuous(n.breaks = 5, labels = selected_labels) +
  labs(x = "Survey Month",
        y = "Plant Growth Index (PGI)")+
  scale_color_manual(values = c("brown1", "deepskyblue"),
                      name = "Irrigation Level",
                     labels = c("High", "Low"))
```
# Mean PGI by Species and Irrigation Level

## Overall

```{r}
# mean PGI by irrigation level
pet_summary_irr <- pet_data_clean %>% 
group_by(irrigation) %>% 
  summarize(mean_growth = mean(growth_rate),
        sd_growth = sd(growth_rate),
        var_growth = var(growth_rate),
        count_growth = length(growth_rate), 
        se_growth = sd_growth/sqrt(count_growth),
        margin_growth = qt(0.95, df = count_growth - 1) * se_growth)

# plotting relationship
ggplot(data = pet_summary_irr, aes (y = mean_growth, x = irrigation, color = irrigation)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Irrigation Level",
        title = "All Species") +
   scale_y_continuous(name = "Mean Plant Growth Index", limits = c(20, 40))+
   theme_bw()
```


Not very clear, let's isolate by species.

```{r}
# isolate species data
 arcto <- pet_data_clean %>% 
   filter(species %in% "arcto")
 
 arte <- pet_data_clean %>% 
   filter(species %in% "arte")
 
 ceo <- pet_data_clean %>% 
   filter(species %in% "ceo")
 
 dip <- pet_data_clean %>% 
   filter(species %in% "dip")
 
 salvia <- pet_data_clean %>% 
   filter(species %in% "salvia")
```

## Arctostaphylos Mean Plant Growth Index by Pot Size and Irrigation
```{r}
# isolated arcto mean PGI by irrigation
arcto_irrigation <- arcto %>% 
   group_by(irrigation) %>% 
   summarize(mean_growth = mean(growth_rate),
             sd_growth = sd(growth_rate),
             var_growth = var(growth_rate),
             count_growth = length(growth_rate),
             se_growth = sd_growth/sqrt(count_growth),
             margin_growth = qt(0.95, df = count_growth - 1) * se_growth)
 
 # plotting relationship
 ggplot(data = arcto_irrigation, aes (y = mean_growth, x = irrigation, color = irrigation)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Irrigation Level",
        y = "Mean PGI",
        title = "Arctostaphylos") +
   theme_bw()
```


```{r}
# isolated arcto mean PGI by pot size
arcto_size <- arcto %>% 
   group_by(size) %>% 
   summarize(mean_growth = mean(growth_rate),
             sd_growth = sd(growth_rate),
             var_growth = var(growth_rate),
             count_growth = length(growth_rate),
             se_growth = sd_growth/sqrt(count_growth),
             margin_growth = qt(0.95, df = count_growth - 1) * se_growth)
 
 # plotting relationship
 ggplot(data = arcto_size, aes (y = mean_growth, x = size, color = size)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Pot Size",
        y = "Mean PGI",
        title = "Arctostaphylos") +
   theme_bw()
```

```{r}
# anovas for arcto size and irrigation
 arcto_irr_anova <- aov(growth_rate ~ irrigation, data = arcto)
 
 arcto_size_anova <- aov(growth_rate ~ size, data = arcto)

summary(arcto_irr_anova)
summary(arcto_size_anova)
```

We can see that the Acrtostaphylos in the low irrigation plot had a significantly higher mean plant growth index than those in the high irrigation plot. The Acrtostaphylos which started in 1 gallon pots had a significantly higher mean plant growth index than those started in 4 inch pots.

## Artemisia Mean Plant Growth Index by Pot Size and Irrigation
```{r}
# isolated arte mean PGI by irrigation
arte_irrigation <- arte %>% 
   group_by(irrigation) %>% 
   summarize(mean_growth = mean(growth_rate),
             sd_growth = sd(growth_rate),
             var_growth = var(growth_rate),
             count_growth = length(growth_rate),
             se_growth = sd_growth/sqrt(count_growth),
             margin_growth = qt(0.95, df = count_growth - 1) * se_growth)
 
 # plotting relationship
 ggplot(data = arte_irrigation, aes (y = mean_growth, x = irrigation, color = irrigation)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Irrigation Level",
        y = "Mean PGI",
        title = "Artemisia") +
   theme_bw()
```


```{r}
# isolated arte mean PGI by pot size
arte_size <- arte %>% 
   group_by(size) %>% 
   summarize(mean_growth = mean(growth_rate),
             sd_growth = sd(growth_rate),
             var_growth = var(growth_rate),
             count_growth = length(growth_rate),
             se_growth = sd_growth/sqrt(count_growth),
             margin_growth = qt(0.95, df = count_growth - 1) * se_growth)
 
 # plotting relationship
 ggplot(data = arte_size, aes (y = mean_growth, x = size, color = size)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Pot Size",
        y = "Mean PGI",
        title = "Artemisia") +
   theme_bw()
```

```{r}
# anovas for arte size and irrigation
 arte_irr_anova <- aov(growth_rate ~ irrigation, data = arte)
 
 arte_size_anova <- aov(growth_rate ~ size, data = arte)

summary(arte_irr_anova)
summary(arte_size_anova)
```
We can see there is no significant difference in the mean growth rate between pot size or irrigation level for Artemisia. The low irrigation and the 1 gallon plants had slightly (but not significantly) higher mean PG.

## Ceanothus Mean Plant Growth Index by Pot Size and Irrigation
```{r}
# isolated ceo mean PGI by irrigation
ceo_irrigation <- ceo %>% 
   group_by(irrigation) %>% 
   summarize(mean_growth = mean(growth_rate),
             sd_growth = sd(growth_rate),
             var_growth = var(growth_rate),
             count_growth = length(growth_rate),
             se_growth = sd_growth/sqrt(count_growth),
             margin_growth = qt(0.95, df = count_growth - 1) * se_growth)
 
 # plotting relationship
 ggplot(data = ceo_irrigation, aes (y = mean_growth, x = irrigation, color = irrigation)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Irrigation Level",
        y = "Mean PGI",
        title = "Ceanothus") +
   theme_bw()
```


```{r}
# isolated arte mean PGI by pot size
ceo_size <- ceo %>% 
   group_by(size) %>% 
   summarize(mean_growth = mean(growth_rate),
             sd_growth = sd(growth_rate),
             var_growth = var(growth_rate),
             count_growth = length(growth_rate),
             se_growth = sd_growth/sqrt(count_growth),
             margin_growth = qt(0.95, df = count_growth - 1) * se_growth)
 
 # plotting relationship
 ggplot(data = ceo_size, aes (y = mean_growth, x = size, color = size)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Pot Size",
        y = "Mean PGI",
        title = "Ceanothus") +
   theme_bw()
```

```{r}
# anovas for ceo size and irrigation
 ceo_irr_anova <- aov(growth_rate ~ irrigation, data = ceo)
 
 ceo_size_anova <- aov(growth_rate ~ size, data = ceo)

summary(ceo_irr_anova)
summary(ceo_size_anova)
```

We can see that the Ceanothus in the high irrigation plot had a significantly larger mean plant growth index than those in the low irrigation plot. There was no significant difference between Ceanothus in 1 gallon and 4 inch pots, however the 1 gallon Ceanothus mean PGI was slightly higher.

## Diplacus Mean Plant Growth Index by Pot Size and Irrigation
```{r}
# isolated dip mean PGI by irrigation
dip_irrigation <- dip %>% 
   group_by(irrigation) %>% 
   summarize(mean_growth = mean(growth_rate),
             sd_growth = sd(growth_rate),
             var_growth = var(growth_rate),
             count_growth = length(growth_rate),
             se_growth = sd_growth/sqrt(count_growth),
             margin_growth = qt(0.95, df = count_growth - 1) * se_growth)
 
 # plotting relationship
 ggplot(data = dip_irrigation, aes (y = mean_growth, x = irrigation, color = irrigation)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Irrigation Level",
        y = "Mean PGI",
        title = "Diplacus") +
   theme_bw()
```


```{r}
# isolated dip mean PGI by pot size
dip_size <- dip %>% 
   group_by(size) %>% 
   summarize(mean_growth = mean(growth_rate),
             sd_growth = sd(growth_rate),
             var_growth = var(growth_rate),
             count_growth = length(growth_rate),
             se_growth = sd_growth/sqrt(count_growth),
             margin_growth = qt(0.95, df = count_growth - 1) * se_growth)
 
 # plotting relationship
 ggplot(data = dip_size, aes (y = mean_growth, x = size, color = size)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Pot Size",
        y = "Mean PGI",
        title = "Diplacus") +
   theme_bw()
```

```{r}
# anovas for dip size and irrigation
 dip_irrigation_anova <- aov(growth_rate ~ irrigation, data = dip)
 
 dip_size_anova <- aov(growth_rate ~ size, data = dip)

summary(dip_irrigation_anova)
summary(dip_size_anova)
```

We can see that there is no significant difference in the mean PGI of Diplacus planted in the low and high level irrigation plots, or between those started in 1 gallon and 4 inch pots.

## Salvia Mean Plant Growth Index by Pot Size and Irrigation
```{r}
salvia_irrigation <- salvia %>% 
   group_by(irrigation) %>% 
   summarize(mean_growth = mean(growth_rate),
             sd_growth = sd(growth_rate),
             var_growth = var(growth_rate),
             count_growth = length(growth_rate),
             se_growth = sd_growth/sqrt(count_growth),
             margin_growth = qt(0.95, df = count_growth - 1) * se_growth)
 
 # plotting relationship
 ggplot(data = salvia_irrigation, aes (y = mean_growth, x = irrigation, color = irrigation)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Irrigation Level",
        y = "Mean PGI",
        title = "Salvia") +
   theme_bw()
```


```{r}
salvia_size <- salvia %>% 
   group_by(size) %>% 
   summarize(mean_growth = mean(growth_rate),
             sd_growth = sd(growth_rate),
             var_growth = var(growth_rate),
             count_growth = length(growth_rate),
             se_growth = sd_growth/sqrt(count_growth),
             margin_growth = qt(0.95, df = count_growth - 1) * se_growth)
 
 # plotting relationship
 ggplot(data = salvia_size, aes (y = mean_growth, x = size, color = size)) +
   geom_point() +
   geom_linerange(aes(ymin = mean_growth - margin_growth, ymax = mean_growth + margin_growth), linewidth = 0.5) +
   labs(x = "Pot Size",
        y = "Mean PGI",
        title = "Salvia") +
   theme_bw()
```

```{r}
# anovas for salvia size and irrigation
 salvia_irrigation_anova <- aov(growth_rate ~ irrigation, data = salvia)
 
 salvia_size_anova <- aov(growth_rate ~ size, data = salvia)

summary(salvia_irrigation_anova)
summary(salvia_size_anova)
```

We can see there is no significant difference in the mean growth rate between pot size or irrigation level for Salvia. The high irrigation and the 4 inch plants had slightly (but not significantly) higher mean PG.


The mean of the plant growth index flattens the data and removes the seasonal oscillations inherent this time-series data. We can also look at the growth rate of each treatment by taking the slope of the line representing the PGI over time. This serves to visualize the inches of growth per month for each species.


# Growth Rate of Species (PGI/time)

## Arctostaphylos Growth Rate

```{r}
irr_labels <- labeller(
  irrigation = c(
    high = "High Irrigation",
    low = "Low Irrigation"))
```

```{r}
#creating a plot of growth rate by irrigation
# split by irrigation
ggplot(data = arcto, aes(x = surv_month, y = growth_rate)) +
  geom_jitter(aes(color = size), na.rm = TRUE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+
  labs(title = "Arctostaphylos")+
  facet_wrap(~irrigation, scales = "free", labeller = irr_labels) +
    scale_x_continuous(name = "Survey Month", limits = c(1, 24)) +
  scale_y_continuous(name = "Plant Growth Index", limits = c(1, 25))+
   scale_color_manual(values = c("brown1", "deepskyblue"),
                      name = " ",
                     labels = c("1 gallon", "4 inch")) +  
  # adding slope lines
  # high irr 1 gal
  geom_abline(data = subset(arcto, irrigation == "high"), 
              aes(slope=0.261, intercept=10.056), 
              color = "brown1") +
  # high irr 4 in
  geom_abline(data = subset(arcto, irrigation == "high"), 
              aes(slope=0.284, intercept=9.285), 
              color = "deepskyblue")+
  # low irr 4 in
   geom_abline(data = subset(arcto, irrigation == "low"), 
              aes(slope=0.382, intercept=8.794), 
              color = "deepskyblue")+
  # low irr 1 gal
  geom_abline(data = subset(arcto, irrigation == "low"), 
              aes(slope=0.554, intercept=10.385), 
              color = "brown1")+
  # adding slope labels
  # low irr 1 gal
 geom_text(data = subset(arcto, irrigation == "low"), 
            aes(x = 12, y = 19, label = "0.554"), 
            color = "brown1", size = 4, fontface = "bold")+
   # low irr 4 in
   geom_text(data = subset(arcto, irrigation == "low"), 
            aes(x = 17, y = 13, label = "0.382"), 
            color = "deepskyblue", size = 4, fontface = "bold")+
    # high irr 1 gal
   geom_text(data = subset(arcto, irrigation == "high"), 
            aes(x = 15, y = 15.5, label = "0.261"), 
            color = "brown1", size = 4, fontface = "bold") +
    # high irr 4 in
  geom_text(data = subset(arcto, irrigation == "high"), 
            aes(x = 17, y = 13, label = "0.284"), 
            color = "deepskyblue", size = 4, fontface = "bold")
```

We can see that the plants in the low irrigation plot had the highest growth rate (PGI/time). 

## Artemisia Growth Rate

```{r}
#creating a plot of growth rate by irrigation
# split by irrigation
ggplot(data = arte, aes(x = surv_month, y = growth_rate)) +
  geom_jitter(aes(color = size), na.rm = TRUE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+
  labs(title = "Artemisia")+
  facet_wrap(~irrigation, scales = "free", labeller = irr_labels) +
    scale_x_continuous(name = "Survey Month", limits = c(1, 24)) +
  scale_y_continuous(name = "Plant Growth Index", limits = c(1, 65))+
   scale_color_manual(values = c("brown1", "deepskyblue"),
                      name = " ",
                     labels = c("1 gallon", "4 inch"))+
 # adding slope lines
  # high irr 1 gal
  geom_abline(data = subset(arte, irrigation == "high"), 
              aes(slope= 1.68335, intercept= 17.28893), 
              color = "brown1") +
  # high irr 4 in
  geom_abline(data = subset(arte, irrigation == "high"), 
              aes(slope = 1.664921, intercept= 10.83617), 
              color = "deepskyblue")+
  # low irr 4 in
   geom_abline(data = subset(arte, irrigation == "low"), 
              aes(slope = 1.992367, intercept= 11.42986), 
              color = "deepskyblue")+
  # low irr 1 gal
  geom_abline(data = subset(arte, irrigation == "low"), 
              aes(slope = 1.940563, intercept= 13.28498), 
              color = "brown1") +
   # adding slope labels
  # low irr 1 gal
 geom_text(data = subset(arcto, irrigation == "low"), 
            aes(x = 19, y = 58, label = "1.941"), 
            color = "brown1", size = 4, fontface = "bold")+
   # low irr 4 in
   geom_text(data = subset(arte, irrigation == "low"), 
            aes(x = 17, y = 35, label = "1.992"), 
            color = "deepskyblue", size = 4, fontface = "bold")+
    # high irr 1 gal
   geom_text(data = subset(arte, irrigation == "high"), 
            aes(x = 18, y = 55, label = "1.683"), 
            color = "brown1", size = 4, fontface = "bold") +
    # high irr 4 in
  geom_text(data = subset(arte, irrigation == "high"), 
            aes(x = 17, y = 33, label = "1.665"), 
            color = "deepskyblue", size = 4, fontface = "bold")
```
We can see that the plants in the low irrigation plot had a higher growth rate than those in the high irrigation plot. It's also interesting to note that the growth rates of the pot sizes within each irrigation level are quite similar to each other, indicating that pot size had little effect on the growth rate of Artemisia.

## Ceanothus Growth Rate

```{r}
#creating a plot of growth rate by irrigation
# split by irrigation
ggplot(data = ceo, aes(x = surv_month, y = growth_rate)) +
  geom_jitter(aes(color = size), na.rm = TRUE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+
  labs(title = "Ceanothus")+
  facet_wrap(~irrigation, scales = "free", labeller = irr_labels) +
    scale_x_continuous(name = "Survey Month", limits = c(1, 24)) +
  scale_y_continuous(name = "Plant Growth Index", limits = c(1, 40))+
   scale_color_manual(values = c("brown1", "deepskyblue"),
                      name = " ",
                     labels = c("1 gallon", "4 inch"))+
 # adding slope lines
  # high irr 1 gal
  geom_abline(data = subset(ceo, irrigation == "high"), 
              aes(slope= 0.858943, intercept= 16.40464), 
              color = "brown1") +
  # high irr 4 in
  geom_abline(data = subset(ceo, irrigation == "high"), 
              aes(slope = 0.929298, intercept= 15.07559), 
              color = "deepskyblue")+
  # low irr 4 in
   geom_abline(data = subset(ceo, irrigation == "low"), 
              aes(slope = 0.766181, intercept= 11.758), 
              color = "deepskyblue")+
  # low irr 1 gal
  geom_abline(data = subset(arcto, irrigation == "low"), 
              aes(slope = 0.999457, intercept= 12.89891), 
              color = "brown1")+
  # adding slope labels
  # low irr 1 gal
 geom_text(data = subset(ceo, irrigation == "low"), 
            aes(x = 19, y = 36, label = "0.999"), 
            color = "brown1", size = 4, fontface = "bold")+
   # low irr 4 in
   geom_text(data = subset(ceo, irrigation == "low"), 
            aes(x = 17, y = 22, label = "0.766"), 
            color = "deepskyblue", size = 4, fontface = "bold")+
    # high irr 1 gal
   geom_text(data = subset(ceo, irrigation == "high"), 
            aes(x = 19, y = 36, label = "0.859"), 
            color = "brown1", size = 4, fontface = "bold") +
    # high irr 4 in
  geom_text(data = subset(ceo, irrigation == "high"), 
            aes(x = 17, y = 27, label = "0.929"), 
            color = "deepskyblue", size = 4, fontface = "bold")
```

We can see that there is more variation in the Ceanothus growth rate, with the 1 gallon Ceanothus in the low irrigation plot having the highest growth rate while the 4 inch Ceanothus in the low irrigation plot have the lowest growth rate. This variation may show that 1 gallon plants preform better, especially when there is less irrigation, or it could be error due to the fact that predation on Ceanothus by deer was quite high. 

## Diplacus Growth Rate
```{r}
#creating a plot of growth rate by irrigation
# split by irrigation
ggplot(data =  dip, aes(x = surv_month, y = growth_rate)) +
  geom_jitter(aes(color = size), na.rm = TRUE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+
  labs(title = "Diplacus")+
  facet_wrap(~irrigation, scales = "free", labeller = irr_labels) +
    scale_x_continuous(name = "Survey Month", limits = c(1, 24)) +
  scale_y_continuous(name = "Plant Growth Index", limits = c(1, 45))+
   scale_color_manual(values = c("brown1", "deepskyblue"),
                      name = " ",
                     labels = c("1 gallon", "4 inch"))+
  # adding slope lines
  # high irr 1 gal
  geom_abline(data = subset(dip, irrigation == "high"), 
              aes(slope= 0.328162, intercept= 19.05336), 
              color = "brown1") +
  # high irr 4 in
  geom_abline(data = subset(dip, irrigation == "high"), 
              aes(slope = 0.063661067, intercept= 22.19639), 
              color = "deepskyblue")+
  # low irr 4 in
   geom_abline(data = subset(dip, irrigation == "low"), 
              aes(slope = 1.258064, intercept= 10.12307), 
              color = "deepskyblue")+
  # low irr 1 gal
  geom_abline(data = subset(dip, irrigation == "low"), 
              aes(slope = 0.731441, intercept= 14.17515), 
              color = "brown1")+
  # adding slope labels
  # low irr 1 gal
 geom_text(data = subset(dip, irrigation == "low"), 
            aes(x = 17, y = 23, label = "0.731"), 
            color = "brown1", size = 4, fontface = "bold")+
   # low irr 4 in
   geom_text(data = subset(dip, irrigation == "low"), 
            aes(x = 20, y = 40, label = "1.258"), 
            color = "deepskyblue", size = 4, fontface = "bold")+
    # high irr 1 gal
   geom_text(data = subset(dip, irrigation == "high"), 
            aes(x = 20, y = 30, label = "0.328"), 
            color = "brown1", size = 4, fontface = "bold") +
    # high irr 4 in
  geom_text(data = subset(dip, irrigation == "high"), 
            aes(x = 12, y = 20, label = "0.064"), 
            color = "deepskyblue", size = 4, fontface = "bold")
```

We can see that the Diplacus in the low irrigation plot had a much higher growth rate than those in the high irrigation plot. In the low irrigation we can see that the 4 inch plants outperformed the 1 gallon plants, however the inverse is true in the high irrigation plot. This further confirms that initial pot size does not have a predictable effect on the plant's growth. 

## Salvia Growth Rate

```{r}
#creating a plot of growth rate by irrigation
# split by irrigation
ggplot(data =  salvia, aes(x = surv_month, y = growth_rate)) +
  geom_jitter(aes(color = size), na.rm = TRUE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))+
  labs(title = "Salvia")+
  facet_wrap(~irrigation, scales = "free", labeller = irr_labels) +
    scale_x_continuous(name = "Survey Month", limits = c(1, 24)) +
  scale_y_continuous(name = "Plant Growth Index", limits = c(1, 65))+
   scale_color_manual(values = c("brown1", "deepskyblue"),
                      name = " ",
                     labels = c("1 gallon", "4 inch"))+
  # adding slope lines
  # high irr 1 gal
  geom_abline(data = subset(salvia, irrigation == "high"), 
              aes(slope= 1.432238, intercept= 25.03053), 
              color = "brown1") +
  # high irr 4 in
  geom_abline(data = subset(salvia, irrigation == "high"), 
              aes(slope = 1.312352, intercept= 18.73221), 
              color = "deepskyblue")+
  # low irr 4 in
   geom_abline(data = subset(salvia, irrigation == "low"), 
              aes(slope = 2.250896, intercept= 7.425395), 
              color = "deepskyblue")+
  # low irr 1 gal
  geom_abline(data = subset(salvia, irrigation == "low"), 
              aes(slope = 1.873007401, intercept= 11.52011), 
              color = "brown1")+
   # adding slope labels
  # low irr 1 gal
 geom_text(data = subset(salvia, irrigation == "low"), 
            aes(x = 17, y = 36, label = "1.873"), 
            color = "brown1", size = 4, fontface = "bold")+
   # low irr 4 in
   geom_text(data = subset(salvia, irrigation == "low"), 
            aes(x = 18, y = 56, label = "2.251"), 
            color = "deepskyblue", size = 4, fontface = "bold")+
    # high irr 1 gal
   geom_text(data = subset(salvia, irrigation == "high"), 
            aes(x = 17, y = 55, label = "1.432"), 
            color = "brown1", size = 4, fontface = "bold") +
    # high irr 4 in
  geom_text(data = subset(salvia, irrigation == "high"), 
            aes(x = 15, y = 32, label = "1.312"), 
            color = "deepskyblue", size = 4, fontface = "bold")
```

We can see that the Salvia in the low irrigation plot had a higher growth rate than those in the high irrigation plot. 

# find error in slope , find diff between averages, can report how much more low grew on avg

# Appearance score
```{r}
# plot to visualize appearance scores
ggplot(data = appearance, aes(x = surv_month, y = score, color = irrigation)) +
       geom_jitter(na.rm = TRUE) +
  theme_bw()+
  facet_wrap(~species + size, labeller = sp_labels) +
  scale_x_continuous(n.breaks = 5, labels = selected_labels) +
  labs(x = "Survey Month",
        y = "Appearance Score",
       title = "Plant Appearance Over Time")+
  scale_color_manual(values = c("brown1", "deepskyblue"),
                      name = "Irrigation Level",
                     labels = c("High", "Low"))
```

# total deaths/mortality in each treatment

```{r}
# plot to visualize appearance scores
ggplot(data = mortality, aes(x = mortality, color = irrigation)) +
       geom_bar(na.rm = TRUE) +
  theme_bw()+
  facet_wrap(~species + size, labeller = sp_labels)  +
  labs(x = "Survey Month",
        y = "Number of Mortalities",
       title = "Total Plant Mortalities")+
  scale_color_manual(values = c("brown1", "deepskyblue"),
                      name = "Irrigation Level",
                     labels = c("High", "Low"))
```




